\section{Introduction and reader's guide}
\label{sec:introduction}

Hexo is played on an unbounded hexagonal grid.  A player wins by occupying
six consecutive cells on one of the three grid axes.  After the opening move,
each turn normally contains two placements.  A move is legal when it is close
to some stone already on the board.  The legal region therefore grows with
the position.  A distant placement can matter later even when it does not
touch a current threat.

This paper studies the defender nodes of threat-space search.  Threat-space
search follows forcing attacking moves and asks whether every defense still
leads to a win.  The natural OR nodes are often narrow because the search has
chosen one attacking continuation.  The AND nodes are wide.  A midgame node
can have hundreds of legal defender replies.  Testing all of them at every
depth can dominate the search even when the tactical line itself is short.
This is the main branching problem addressed here.

The usual local shortcut is unsafe.  A defender stone outside the visible
threats can expand the radius-eight legality frontier.  Later stones can use
that new support to reach a cell required by the attacking line.  The same
remote play can also start a run that completes before the planned attack.
A sound restriction must account for occupancy, the windows changed by a
stone, and the legality frontier created by that stone.  A rule based only on
the current threat cells misses two of those effects.

The remedy is a certificate.  A certificate is a finite attacking strategy
tree with exact positions, legal transitions, terminal witnesses, and a
declared resolution time.  It also records which future cells must remain
available.  A separate verifier checks the tree.  Search may use aggressive
move ordering or incomplete candidate generation, but a claimed win is
accepted only when the certificate covers every omitted defender reply.
This separates fast discovery from trusted verification.

The central result gives a sound coverage test for each ordinary defender
node.  Its mandatory zone has four parts.  Direct obligations protect cells
that are already legal.  Ranked seed guards stop a remote reply from opening a
chain into an obligation before its deadline.  Touched-window guards protect
windows that already contain a defender stone and can complete in time.
Virgin-window guards cover a legal seed that can reach and fill an empty
window before the attack resolves.  Each obligation and each window uses a
local clock.  A deeper certificate can therefore protect more territory than
a shallow certificate, while unrelated long branches do not inflate every
node.

The proof compares a real play with a ghost walk through the certificate.  A
searched defender reply is copied.  An omitted reply is placed only on the
real board, while a searched filler is placed on the ghost board.  The two
positions can then differ in defender stones but continue to share attacker
stones, ply, mover, and budget.  The four-part zone prevents those differences
from occupying a required attacker cell or completing a defender window
before the mapped resolution.  Terminal witnesses then transfer from the
ghost path to the real play.

Some forcing nodes permit a smaller hazard count.  If the current threats
force the defender to choose from a tight hitting kernel, an exact copied hit
does not create a real-only legality frontier.  The gate calculus debits that
hit from the role clock and charges it to a particular defender window only
when the hit lies in that window.  An ignored gate keeps the full cost of the
remaining defender turn, because the defender may try to win before the
attacker uses a surviving threat.  This is a proved, target-specific debit.
It does not reduce the general path budget, assign zero cost to a substituted
reply, or guarantee that every final searched set becomes smaller.

Several finite results sharpen the certificate format.  A LOSS leaf needs at
most three threat windows when one defender placement remains and at most five
when two remain.  Both caps are attained by Hexo configurations.  A separate
kernel theorem handles fully forced regions entered with identical real and
ghost positions.  Transition envelopes allow a dismissal to select a
particular searched child, but they must include the current transition.
Finite acyclic certificate graphs are sound when shared labels and clocks are
consistent.

The pairing result locates an exact lattice threshold.  No position-independent
matching can put a defensive pair inside every length-six Hexo window.  The
density obstruction becomes equality at length seven, where an index-twelve
periodic perfect matching covers every window exactly once.  The construction
is minimal among periodic constructions.  This closes the global static
pairing question for the Hexo board while leaving position-dependent local
pairings available as search devices.

The potential layer describes a different defensive certificate.  A weighted
sum over attacker-touched live windows can block a fixed family forever when
the defender moves first and the sum is below one.  The same scalar applied to
the changing family protects a finite horizon, not an unrestricted game.
Newly touched windows add mass.  An exhaustive 124-branch calculation shows
that every dynamic greedy tie-breaking rule can lose from a position below
the threshold.  A short clean-escape proof also shows that no defense can
preserve that raw sublevel set at every turn.  Finite-horizon extensions and a
finite reduction remain valid.  The existence of a different non-greedy
forever defense is open.

Local domination gives three additional solver rules.  A dead-cell reply can
be dismissed when its named substitute has no new legality frontier.  Two
hits of one threat can be interchangeable under dead-spoke and support
conditions.  Two placements of the same turn commute when both were legal at
turn start and neither first placement wins.  These rules are independent of
the ranked zone and can reduce branching before certificate verification.

The limit map separates absolute game constants from bounds that are final
only inside a named certificate or potential framework.  It records which
thresholds are attained, which are merely sufficient, and which questions
remain open.  This prevents a local verifier bound from being read as a
position-independent law of the game.

\begin{figure}[tbp]
  \centering
  \begin{tikzpicture}
    \begin{scope}[xshift=-7.0cm,scale=.39]
      \HexBoardRange{0}{0}{3}
      \HexAttackerList{-2/0,-1/0,0/0,1/0}
      \HexDefenderList{-2/2,2/-2}
      \HexShadeList[zone direct]{2/0,3/0}
      \node[font=\scriptsize,align=center] at (0,-7.0)
        {one tactical position\\many legal replies};
    \end{scope}
    \draw[->,>=Stealth,line width=.8pt] (-4.35,0) -- (-2.75,0);
    \node[font=\scriptsize,above] at (-3.55,0) {search};
    \begin{scope}[xshift=-.65cm,yshift=.35cm]
      \node[cert node] (r) {exact AND node};
      \node[cert node,below left=7mm and 5mm of r] (c1) {searched\\reply};
      \node[cert node,below right=7mm and 5mm of r] (c2) {searched\\reply};
      \draw[cert edge] (r)--(c1);
      \draw[cert edge] (r)--(c2);
      \node[font=\scriptsize,align=center] at (0,-2.25)
        {finite certificate\\with clocks and witnesses};
    \end{scope}
    \draw[->,>=Stealth,line width=.8pt] (1.55,0) -- (3.05,0);
    \node[font=\scriptsize,above] at (2.3,0) {verify};
    \begin{scope}[xshift=5.2cm]
      \node[draw,text width=3.35cm,minimum height=2.45cm,align=left,
        font=\scriptsize] (z) {direct obligations\\ranked seed guards\\
        touched-window guards\\virgin-window guards};
      \node[font=\scriptsize,align=center,below=3mm of z]
        {verified coverage of every omission};
    \end{scope}
  \end{tikzpicture}
  \caption{Orientation.  Search proposes a finite attacking certificate.
  Verification reconstructs the obligations and clocks that justify the
  omitted defender replies.}
  \label{fig:orientation}
\end{figure}

\subsection{How to read the paper}

The formal development starts in \cref{sec:game}.  The companion directory
contains the longer derivations and the records of finite checks
\cite{hexobot-companions}.  The paper gives complete proofs for the dismissal
theorem, the ranked-zone reading, the gate soundness theorem, the sparse
witness theorem, the pairing threshold, and the short boundary results proved
here.  Other supporting statements are given with proof
sketches and a companion citation.

\begin{table}[tbp]
  \centering
  \small
  \begin{tabular}{@{}p{.20\linewidth}p{.31\linewidth}p{.39\linewidth}@{}}
    \toprule
    Purpose & Main route & Optional material on a first pass \\
    \midrule
    Solver implementation &
      \cref{sec:game,sec:certificates,sec:main-theorem,sec:computable-zones,sec:gates,sec:validation,sec:conclusions} &
      The density proof in \cref{sec:pairing} and the detailed potential
      boundary in \cref{sec:potential}. \\
    Combinatorial results &
      \cref{sec:geometry,sec:witnesses,sec:pairing,sec:potential,sec:tightness} &
      The verifier data structures in \cref{sec:certificates,sec:computable-zones}. \\
    General orientation &
      This section, \cref{sec:game,sec:main-theorem,sec:gates,sec:conclusions} &
      Proof details in \cref{sec:certificates,sec:potential} and the
      crosswalk in \cref{sec:appendix}. \\
    \bottomrule
  \end{tabular}
  \caption{Reader routes through the paper.}
  \label{tab:reader-routes}
\end{table}

Threat-space search and df-pn supply the algorithmic setting
\cite{allis-1994,nagai-2002}.  The conclusions return to implementation and
state which tests are safe, which reductions are conditional on extra
certificate data, and which questions remain heuristic.
