\section{Locality and elementary geometry}
\label{sec:geometry}

The certificate proof uses two kinds of locality.  A six-cell window has
small diameter, while a new stone changes legality only within radius eight.
The first group of statements controls tactics inside one window.  The second
group lists every channel that must be checked when two positions differ.
Complete derivations are recorded in the companion proof
\cite{hexobot-companions}.

A line segment with many stones has few empty cells.  The farthest empty
cannot be farther from all stones than the number of empties permits.

\begin{lemma}[Packing bound]
\label{lem:L1}
In a window with $k\ge1$ stones of one colour and no stones of the other,
every empty cell is within distance $6-k$ of a same-window stone.  The bound
is tight.
\end{lemma}

\begin{proof}[Proof sketch]
Index the window cells from $0$ to $5$ along its axis.  If an empty cell has
nearest in-window stone at axis distance $\delta$, then that cell and the
$\delta-1$ cells between them are empty.  Hence $\delta\le6-k$.  Axis distance
equals hex distance inside a window by \cref{fact:F1}.  Stones at offsets
$6-k,\ldots,5$ and an empty at offset $0$ attain equality.
\end{proof}

\begin{plainterms}
Four stones in a live window put every remaining cell within distance two.
Three stones give distance three.  These two substitutions drive most of the
local tactical bounds.
\end{plainterms}

\begin{figure}[tbp]
  \centering
  \begin{subfigure}[t]{.47\textwidth}
    \centering
    \begin{tikzpicture}[scale=.48]
      \HexPatch{0/0,1/0,2/0,3/0,4/0,5/0}
      \HexWindow[window one]{0}{0}{1}{0}
      \HexAttackerList{3/0,4/0,5/0}
      \HexMark{0}{0}{cross}{}
      \HexDistance{0}{0}{3}{0}{$6-k=3$}
    \end{tikzpicture}
    \caption{The equality case for $k=3$.}
  \end{subfigure}
  \hfill
  \begin{subfigure}[t]{.47\textwidth}
    \centering
    \begin{tikzpicture}[scale=.48]
      \HexPatch{0/0,1/0,2/0,3/0,4/0,5/0}
      \HexWindow[window two]{0}{0}{1}{0}
      \HexAttackerList{2/0,3/0,4/0,5/0}
      \HexMark{0}{0}{cross}{}
      \HexDistance{0}{0}{2}{0}{$6-k=2$}
    \end{tikzpicture}
    \caption{The threat case for $k=4$.}
  \end{subfigure}
  \caption{The packing bound and its tight configurations.  A cross marks the
  empty cell at offset zero.}
  \label{fig:packing}
\end{figure}

Whenever a window contains at least one stone and at least one empty, the
boundary between occupied and empty offsets gives an adjacent pair.

\begin{lemma}[Boundary pair]
\label{lem:L2}
Any window containing at least one stone and at least one empty cell has an
empty cell at distance one from a same-window stone.
\end{lemma}

\begin{proof}[Proof sketch]
The occupied offsets form a nonempty proper subset of a six-cell line.  At a
boundary of this subset, an occupied offset and an empty offset are adjacent.
Their axis distance and hex distance are one.
\end{proof}

The packing bound places every current tactical answer close to a stone that
already supports the same window.

\begin{theorem}[Single-turn tactical locality]
\label{thm:T1}
Every placement that completes a win within the current turn, and every empty
cell of every threat window, lies within distance two of an existing stone.
\end{theorem}

\begin{proof}[Proof sketch]
A completion using at most two placements starts from an alive window with at
least four stones of that player.  A threat has the same count lower bound by
\cref{def:D5}.  Apply \cref{lem:L1} with $k=4$.
\end{proof}

\begin{plainterms}
Immediate wins and direct threat answers are always legal after the opening.
They lie inside the smaller radius-two neighborhood, even though the game
allows radius-eight placements.
\end{plainterms}

Creating a new threat needs three earlier stones in its witness window.  This
gives the radius used by the common local candidate generator.

\begin{theorem}[Threat-creation locality]
\label{thm:T2}
A placement that creates an $X$-threat lies within distance three of an
existing $X$ stone.  The bound is tight.
\end{theorem}

\begin{proof}[Proof sketch]
Immediately before the placement, its witness window contains at least three
$X$ stones.  Apply \cref{lem:L1} with $k=3$.  Stones at offsets $3,4,5$ and a
placement at offset $0$ attain distance three.
\end{proof}

\begin{plainterms}
Radius three contains every move that turns an existing three-stone support
into a threat.  It does not contain every setup move in a long certificate.
A fourth future setup placement can enter a window that is empty at the
current node.
\end{plainterms}

\subsection{The three causal channels}

A transfer argument compares positions that differ by one or more defender
stones.  The next lemma gives a complete checklist for the one-stone step.

\begin{lemma}[Difference accounting]
\label{lem:L3}
Let $P'=P\mathbin{\pm}\{x\}$ differ from $P$ by one stone at cell $x$.  Every
game-relevant predicate difference is confined to the following channels:
\begin{enumerate}[label=\textup{(C\arabic*)}]
  \item occupancy and ownership of $x$;
  \item the masks of the eighteen windows containing $x$, including every
  completion, alive/dead, threat, $\ownwinnow$, $\hit$, and $\mhs$ predicate;
  \item the legality frontier
  $\{c:\dist(c,x)\le8\}\setminus\operatorname{dom}\sigma$.
\end{enumerate}
\end{lemma}

\begin{proof}[Proof sketch]
The transition rule reads occupancy at the move cell, the legality predicate,
and incident window masks.  All tactical predicates are functions of window
masks.  A stone changes only its eighteen incident masks and supplies legality
only inside its radius-eight ball.  The position has no other mutable state.
\end{proof}

\begin{plainterms}
Matching threat counts is not enough.  A valid simulation must also preserve
the cells needed later and explain any new legal cells.  The main coupling and
the domination patterns use this same checklist.
\end{plainterms}

Stones never leave the board.  This gives the direction in which window states
can change along a play.

\begin{lemma}[Permanence]
\label{lem:L4}
Window masks are coordinatewise nondecreasing along every play.  Dead windows
remain dead.  A window alive for $X$ at ply $t$ is alive for $X$ at every
earlier ply.
\end{lemma}

\begin{proof}[Proof sketch]
Each transition adds one stone and removes none.  A mixed-colour mask cannot
lose either colour.  If an opponent stone had occupied a window earlier, it
would still occupy it at ply $t$.
\end{proof}

Adding defender stones away from named cells and windows is safe in a
particular direction.  The exact completion clause prevents an invalid
monotonicity shortcut.

\begin{lemma}[Inert-extension transfer]
\label{lem:L5}
Let $P'=P+X$, where $X$ is a finite set of defender stones.  Let
$\mathcal W$ be a set of windows and $C$ a set of cells such that no
$x\in X$ lies in a member of $\mathcal W$ and $X\cap C=\varnothing$.  Then:
\begin{enumerate}[label=\textup{(\alph*)}]
  \item every cell of $C$ empty in $P$ is empty in $P'$;
  \item every $W\in\mathcal W$ has identical masks in $P$ and $P'$;
  \item $\Legal(P)\setminus X\subseteq\Legal(P')$, and every cell of $C$
  legal in $P$ is legal in $P'$;
  \item the $\A$-complete windows of $P'$ are exactly the $\A$-complete
  windows of $P$ that avoid $X$.
\end{enumerate}
Thus an attacker completion on a window known to avoid $X$ transfers from
$P$ to $P'$, and every attacker completion in $P'$ was available in $P$.
\end{lemma}

\begin{proof}[Proof sketch]
The first two parts are occupancy and mask equality.  Added stones can add
legality support, while their occupied cells leave the legal set.  A window
completed by the attacker must contain no defender stone, so exactly the
$X$-free completions remain.
\end{proof}

\begin{plainterms}
Extra defender stones can make more moves legal, but they can also block an
attacker window.  Transfer is valid only for named cells and windows that
avoid those stones.
\end{plainterms}

Deleting defender stones helps attacker masks but can remove the only support
that made a later attacker placement legal.

\begin{lemma}[Deletion monotonicity and well-formed legality]
\label{lem:L6}
Let $P''=P-Y$, where $Y$ is a finite set of defender stones of $P$.  Then:
\begin{enumerate}[label=\textup{(\alph*)}]
  \item every attacker threat of $P$ is an attacker threat of $P''$, so the
  defender's $\mhs(P'')\ge\mhs(P)$ and defender $\lambda^1$ LOSS verdicts
  transfer from $P$ to $P''$;
  \item attacker-complete windows of $P$ are attacker-complete in $P''$;
  \item a cell empty in $P$ is empty or newly empty in $P''$;
  \item legality may shrink.  Every relied-upon attacker placement must be
  within distance eight of a stone outside $Y$.  In the uses below it is
  within distance three of an attacker stone, as in \cref{thm:T2}, or within
  distance eight of an unaffected root stone.
\end{enumerate}
\end{lemma}

\begin{proof}[Proof sketch]
Deleting defender stones cannot kill an attacker threat or completion.  It
can only make occupied cells empty.  It can also delete a legality witness,
which is why the last condition is independent of mask monotonicity.
\end{proof}

\begin{plainterms}
Removing defender stones makes the attack no weaker inside existing windows.
It can still make an attacker move illegal.  Every deletion argument must name
an attacker or root stone that continues to justify that move.
\end{plainterms}

The final elementary count separates stones already in a window from defender
placements that remain before a deadline.

\begin{lemma}[Completion counting]
\label{lem:L7}
If $\cnt_\D(W,P)=s$ and the defender makes at most $D$ placements at plies
strictly before $T$, then $W$ cannot become defender-complete before $T$
unless $s+D\ge6$.
\end{lemma}

\begin{proof}[Proof sketch]
A defender completion contains six defender stones.  At most $s$ are present
at $P$, and at most $D$ further stones can be placed before $T$.
\end{proof}

These lemmas do not by themselves define a safe reply set.  The certificate
must say which cells and windows are relevant, and it must attach a deadline
to each one.  That data is introduced next.
