\section{Horizon-parameterized dismissal soundness}
\label{sec:main-theorem}

The proof runs the attacker's real strategy beside a bookkeeping play, called
the ghost, that remains on certificate branches.  The two boards have the
same attacker stones.  The set $X$ contains dismissed defender stones present
only in the real game, while $Y$ contains searched filler stones present only
in the ghost.  The coupling keeps both games at the same ply and budget while
tracking the effects of these two differences.

\begin{figure}[tbp]
  \centering
  \begin{tikzpicture}[scale=.39]
    \begin{scope}[xshift=-8.2cm]
      \HexBoardRange{0}{0}{3}
      \HexAttacker{-1}{0}\HexAttacker{0}{0}\HexAttacker{1}{-1}
      \HexDefender{-1}{1}\HexDefender{2}{-1}
      \node[defender stone,fill=black!30] at \HexAt{1}{1} {};
      \draw[line width=1.2pt] \HexAt{1}{1} ++(-.22,-.22) -- ++(.44,.44);
      \draw[line width=1.2pt] \HexAt{1}{1} ++(-.22,.22) -- ++(.44,-.44);
      \node[font=\small] at (0,-6.8) {real game};
    \end{scope}
    \begin{scope}[xshift=8.2cm]
      \HexBoardRange{0}{0}{3}
      \HexAttacker{-1}{0}\HexAttacker{0}{0}\HexAttacker{1}{-1}
      \HexDefender{-1}{1}\HexDefender{2}{-1}
      \node[defender stone,fill=white] at \HexAt{-2}{0} {};
      \draw[line width=1.15pt] \HexAt{-2}{0} circle (.36);
      \node[font=\small] at (0,-6.8) {ghost game};
    \end{scope}
  \end{tikzpicture}

  \vspace{1ex}
  \begin{tabular}{@{}cl@{\qquad}cl@{}}
    \tikz\node[attacker stone,minimum size=3.7mm] {}; & attacker stone &
    \tikz\node[defender stone,minimum size=3.7mm] {}; & shared defender stone \\
    \tikz{\node[defender stone,fill=black!30,minimum size=3.7mm] {};
      \draw[line width=.8pt] (-.12,-.12)--(.12,.12);
      \draw[line width=.8pt] (-.12,.12)--(.12,-.12);} &
      $X$ (dismissed, real only) &
    \tikz{\node[defender stone,minimum size=3.7mm] {};
      \draw[line width=.8pt] (0,0) circle (.15);} &
      $Y$ (filler, ghost only)
  \end{tabular}
  \caption{The coupled positions have identical attacker stones and shared
  defender stones.  Canonical difference sets record one real-only dismissed
  stone and one ghost-only filler.}
  \label{fig:ghost-coupling}
\end{figure}

The main result turns the local zone conditions into a strategy against every
legal defender reply.

\begin{theorem}[Horizon-parameterized dismissal soundness]
\label{thm:T3}
This statement is proved for valid zone-carrying certificates obeying the
grammar of \cref{def:D9} and satisfying \cref{zone:Z1,zone:Z2,zone:Z4,zone:Z5},
with exact $D_N$ and the full, rather than sharpened, defender-placement
budget.  Let $\mathcal C$ be such a certificate for ``$\A$ wins from $P_0$ by
ply $T$.''  Then $\A$ wins from $P_0$ against every defender strategy over
the full legal move set by a ply at most $T$.  Hence every dismissal in
$\mathcal C$ is sound.
\end{theorem}

\begin{proof}
We construct the attacker's real-game strategy by coupling the real position
$R$, initially $P_0$, to a walk on $\mathcal C$ with ghost position $G$,
also initially $P_0$.  We maintain all invariants of \cref{def:D12}: equal
ply, mover, and budget; identical attacker stones; the canonical defender
differences $X$ and $Y$; and the protection and frontier-clear properties.
Leaves are processed as soon as the walk reaches them.  The induction
parameter is $T$ minus the current ply.

\medskip\noindent
\textbf{Step O (ghost at an OR node, designated placement $c$).}
The cell $c$ is empty in the ghost because the certificate is valid.  It is
not in $X$: it belongs to the core and therefore to the protected set of
every ancestor by \cref{prop:prot-mono}, while the coupling keeps $X$ out of
the protected set.  It is not in $Y$, since cells of $Y$ are ghost defender
stones and $c$ is ghost-empty.  Thus $c$ is empty in the real position.

The cell is also real-legal.  By \cref{zone:Z4}, it lies within distance $8$
of an attacker stone of its predecessor position or a root stone.  These
stones are identical in $R$ and $G$.  The attacker plays $c$ in both games.
The attacker stones remain identical, and $X$ and $Y$ do not change.

If this placement completes the named witness $W^*$ in the ghost, then
$W^*$ belongs to the core.  Hence $X\cap W^*=\varnothing$.  Also
$Y\cap W^*=\varnothing$, because an attacker-complete ghost window contains
no ghost defender stone.  Its masks agree in the two games by
\cref{lem:L5}, so the real placement completes $W^*$ at the same ply, which
is at most $T$.

The real placement may instead complete another window that the ghost does
not.  Such a window is blocked by $Y$ in the ghost but not in the real game,
as characterized by \cref{lem:L5,lem:L6}.  The real game then terminates
immediately with an attacker win before $T$.  If neither completion occurs,
the coupling recurses on the child.

\medskip\noindent
\textbf{Step A (ghost at an AND node $N$, real defender placement $d$).}
The real-legal cell $d$ is either ghost-occupied, ghost-empty and searched,
or ghost-empty and unsearched.  These cases are exhaustive.

\smallskip\noindent
\emph{Filler subroutine.}
This subroutine is used in A2 and A3.  The ghost must consume a defender ply.
Choose any $d_0\in S(N)$.  This set is nonempty by the certificate grammar,
and every member is ghost-legal and ghost-empty.  The ghost plays $d_0$.  If
$d_0$ was real-occupied, then $d_0\in X$ and the update is
$X\leftarrow X\setminus\{d_0\}$: the ghost has caught up.  Otherwise the
update is $Y\leftarrow Y\cup\{d_0\}$.  Recomputing the canonical differences
preserves the coupling invariants.  Every new $Y$-cell is a searched
placement, and $X$ can only shrink.

\smallskip\noindent
\textbf{A1.} Suppose $d$ is ghost-empty and $d\in S(N)$.  Both games play
$d$, their difference sets do not change, and the ghost descends to the
$d$-child.

\smallskip\noindent
\textbf{A2.} Suppose $d$ is ghost-occupied.  It is a ghost defender stone:
ghost attacker stones also occur in the real game, while $d$ is real-empty.
Thus $d\in Y$.  The real game plays $d$, removes it from $Y$, runs the filler
subroutine, and descends to the $d_0$-child.

\smallskip\noindent
\textbf{A3.} Suppose $d$ is ghost-empty and $d\notin S(N)$.  This is a
dismissal.  First, $d\notin\Prot(N)$.  If $d\in\Legal(P_N)$, this follows
from \cref{zone:Z2}, since protected legal cells are searched.  If
$d\notin\Legal(P_N)$ and is real-legal only through the $X$-expanded
frontier, \cref{lem:L9} excludes a protected-not-legal cell.  The assumption
also excludes a protected legal cell.

Second, $d$ is frontier-clear.  If $d\in\Legal(P_N)$, a non-clear cell lies
in the band of \cref{zone:Z5} and would have been searched.  If
$d\notin\Legal(P_N)$, \cref{lem:L9} supplies clearance.  The real game plays
$d$ and updates
\[
  X\leftarrow X\cup\{d\},
  \qquad
  \widehat X\leftarrow\widehat X\cup\{d\}.
\]
The protection and clearance invariants hold.  The ghost runs the filler
subroutine and descends to the $d_0$-child.  This is also where
\cref{lem:L9} prevents a dismissed-stone corridor from reaching protected
frontier territory.

\medskip\noindent
\textbf{Anchored no-completion argument.}
We show that the real defender cannot complete a window before ply $T$.
For every window $W$, the defender stones in the real position are the ghost
defender stones, plus $X$ and minus $Y$.  Consequently,
\begin{equation}
  \cnt_\D(W,R)
  =\cnt_\D(W,G)+|X\cap W|-|Y\cap W|
  \le \cnt_\D(W,G)+|\widehat X\cap W|.
  \tag{MI}\label{eq:MI}
\end{equation}
Suppose that $W$ becomes defender-complete in the real play at a ply less
than $T$.

If no placement in $W$ was ever dismissed, then
$\widehat X\cap W=\varnothing$.  Inequality \eqref{eq:MI} gives
$\cnt_\D(W,G)\ge6$ at that ply.  The ghost follows an actual certificate
line, but the grammar of \cref{def:D9} forbids defender-terminal edges and
requires exact successors.  No valid line can contain this completion, a
contradiction.

Otherwise let $x^*$ be the first cell of $W$ ever placed into
$\widehat X$, and let $N^*$ be its dismissal node.  Write $P(N^*)$,
$R(N^*)$, and $D(N^*)$ for the ghost position, real position, and remaining
defender count at that node.  Step A3 gives
$x^*\notin\Prot(N^*)$.  The window $W$ is alive for the defender at $N^*$:
it completes later, and defender-aliveness persists backward by
\cref{lem:L4}.  The cell $x^*$ is one of its empties.  If
\[
  \cnt_\D(W,P(N^*))+D(N^*)\ge6,
\]
then $W$ qualifies for the completion guard and all its empties, including
$x^*$, lie in $\Prot(N^*)$, a contradiction.  Hence the displayed sum is
less than $6$.

Before $x^*$, no cell of $W$ belongs to $\widehat X$, so \eqref{eq:MI}
gives
\[
  \cnt_\D(W,R(N^*))\le\cnt_\D(W,P(N^*)).
\]
The plays are synchronized.  From $x^*$ onward, at most $D(N^*)$ real
defender placements occur strictly before $T$.  Applying \cref{lem:L7} at
$N^*$ shows that $W$ cannot reach count $6$ before $T$, again a
contradiction.

\medskip\noindent
\textbf{Leaf transfer.}
Post-leaf plies are analyzed directly in the real game; the coupling is not
extended past the leaf.

At a WIN leaf, \cref{def:D9} provides a checked $\ownwinnow$ witness with
count $5$, or with count $4$ and both remaining attacker placements.  No
defender placement is interleaved.  The witness window belongs to the core,
so it is disjoint from $X$; it is defender-free in the ghost and therefore
disjoint from $Y$.  Its masks agree by \cref{lem:L5}.  Its empty cells lie
within distance $2$ of shared attacker stones by \cref{lem:L1}, so they are
legal in the real game.  They are also real-empty: $X$ avoids the core and
$Y$-cells are absent from the real position.  The attacker fills the cells
and completes the window at the declared resolution ply, at most $T$.

At a LOSS leaf with defender budget $b$, the named family $\mathcal T$ has
hitting number greater than $b$.  The real and ghost masks and empty sets of
every $W\in\mathcal T$ are identical at the leaf.  Such windows belong to
the core, so $X$ avoids them.  They are attacker-alive in the ghost, so they
contain no ghost defender stone and in particular no $Y$-cell.  Attacker
stones are shared.  Thus the real hitting number of $\mathcal T$ is the same
as the ghost hitting number and is greater than $b$.

It remains to exclude a real defender $\ownwinnow$ at the leaf.  Let $W$ be
a putative witness.  If $\widehat X\cap W=\varnothing$, the leaf-time form
of \eqref{eq:MI} gives
$\cnt_\D(W,R)\le\cnt_\D(W,G)$.  The attacker mask and budget agree, so the
real predicate would imply the ghost predicate.  This contradicts the
verifier check required at the leaf.  If
$\widehat X\cap W\ne\varnothing$, the second case of the anchored argument,
applied at the first dismissed cell $x^*\in W$, is pure counting over every
real defender placement before $T$.  It does not require the coupling to
continue.  It forbids $W$ from reaching count $6$ before $T$.  Yet a
current-turn win would complete before the declared resolution, which is at
most $T$.  This is also a contradiction.

Now let the real defender play any complete remainder $H$ of the current
turn.  Since $|H|\le b$ is smaller than the hitting number of $\mathcal T$,
some $W_H\in\mathcal T$ satisfies $E(W_H)\cap H=\varnothing$.  The window
remains attacker-alive, because its leaf masks agreed with the ghost and its
cells avoid $H$.  It has at most two empty cells.  By \cref{lem:L1}, they lie
within distance $2$ of shared, permanent attacker stones and remain legal and
empty in the real game.  The attacker completes $W_H$ during the following
turn, using one placement for a count-$5$ witness and two for a count-$4$
witness.  Completion occurs by
\[
  \text{leaf-ply}+b+2\le T.
\]
The set $H$ may contain hits on other witness windows, quiet moves, or
frontier moves.  The argument uses only its cardinality and permanence.
This completes the induction.
\end{proof}
