\section{Computable zones and local reductions}
\label{sec:computable-zones}

The certificate definitions describe finite verifier checks, but a solver
still needs a direct rule for constructing the mandatory searched set.  The
ranked zone theorem gives that rule at an ordinary AND node.  Two narrower
reductions apply when the current threats force a fresh equal-position tree,
or when a dismissal carries a transition-specific substitution envelope.

\subsection{The ranked zone theorem}

The four sets from \cref{def:D11} account for direct occupancy, legality
chains, touched completion windows, and virgin completion windows.  Their
union is the complete ordinary verifier zone.

\begin{theorem}[Ranked zone reading]
\label{thm:T4}
\upshape
At an ordinary AND node $N$, define
\[
 Z(N)=\Zdir(N)\cup\Zseed(N)\cup\Ztouch(N)\cup\Zvirgin(N).
\]
Any independently nonempty searched set $S(N)\supseteq Z(N)$, together with
\cref{zone:Z4}, is sufficient for \cref{thm:T3}.  If $Z(N)$ is empty, the
certificate grammar still requires one arbitrary legal searched fallback.
The uniform admissible assignment $r_N(y)=B(N)$ replaces $\Zseed$ by
\[
 \Legal(P_N)\cap B_{8(B(N)-1)}
 \bigl(\Prot(N)\setminus(\Legal(P_N)\cup\Stones(P_N))\bigr).
\]
Exact cell ranks can only reduce that uniform obligation band.  Exact window
exposures likewise reduce the completion guards obtained from the uniform
$B$ clock.  Larger admissible upper bounds remain sound.  Current
threat-hitting cells are optional search candidates, not terms of $Z(N)$.
\end{theorem}

\begin{plainterms}
At an ordinary node the verifier takes the union of four finite sets.  The
solver may search more cells, but it may not omit a cell of that union unless
it supplies a valid substitution envelope.  An empty union is not an empty
AND node: one legal fallback is still needed to advance the ghost play.
\end{plainterms}

\begin{proof}
Let $S(N)$ be independently nonempty and contain $Z(N)$.  Since it contains
$\Zdir\cup\Ztouch\cup\Zvirgin$, it satisfies \cref{zone:Z2}.  Since it
contains $\Zseed$, it satisfies
\hyperref[zone:Z5prime]{clause \textup{(Z5$'$)}}.  The remaining zone-carrying
condition is the stated well-formed-legality check \cref{zone:Z4}.  The node
therefore satisfies \cref{def:D11}, and \cref{thm:T3} gives pathwise
dismissal soundness.

The empty-union case still falls under \cref{def:D9}, whose AND grammar
requires $S(N)\ne\varnothing$ independently of the coverage clauses.  For
the comparison of clock choices, an exact role rank is bounded by every
admissible uniform rank assigned to that role.  The ball radius
$8(r_N(y)-1)$ is monotone in the rank, so replacing the exact rank by $B(N)$
can only enlarge the seed band.  The recurrences in \cref{def:D14,def:D16}
give $E_N^\D(W)\le B(N)$, and both the touched threshold and virgin radius are
monotone in the exposure.  Thus exact exposure cannot enlarge the
corresponding uniform completion guard.  Conversely, replacing any exact
label by a larger verifier-checked upper bound preserves every inequality in
\cref{lem:L9prime,lem:L12}.  No part of this argument uses the current
threat-hitting set.
\end{proof}

\subsection{The full-clock calculation in the running example}
\label{ex:running-zone-computation}

First read the certificate fragment of \cref{fig:running-certificate} with the
ordinary D11 full clocks, before replacing its root by a forcing gate.  Its
eight future attacker and LOSS-witness carriers are
\[
 \begin{split}
 \Prot(P_0)=\{&(3,20),(0,20),(1,20),(2,20),\\
              &(0,21),(0,22),(1,19),(2,18)\}.
 \end{split}
\]
Every listed cell is legal at $P_0$, and every role has exact root rank one.
It follows that
\[
 \Zdir(P_0)=\Prot(P_0),\qquad \Zseed(P_0)=\varnothing.
\]

Along the sole certificate branch, for any defender-alive queried window $V$,
\[
 E_{P_0}^{\D}(V)=
 \begin{cases}
  1,&V\cap\{(3,20),(0,20)\}\ne\varnothing,\\
  3,&V\cap\{(3,20),(0,20)\}=\varnothing.
 \end{cases}
\]
Enumerating the windows incident with the root defender stones leaves exactly
four touched qualifiers,
\[
 V_j=\{(q,40):j\le q\le j+5\},\qquad j=-3,-2,-1,0.
\]
Each has defender count three and exposure three.  Their empty-cell union is
\[
 \Ztouch(P_0)=
 \{(q,40):q\in\{-3,-2,-1,3,4,5\}\}.
\]
No all-empty window has exposure at least six, so
$\Zvirgin(P_0)=\varnothing$.  Thus the numerical ordinary-zone decomposition
has sizes $8,0,6,0$ in the order direct, seed, touched, virgin.  The actual
root is instead a forcing gate and searches its exact kernel $K=\{h\}$, as
defined in \cref{sec:gates}.

For
\[
 W=\{(q,40):0\le q\le5\},
\]
the root has
\[
 \cnt_\D(W,P_0)=3,
 \qquad h\notin W.
\]
The full D16 exposure counts the exact defender reply at the root and the two
placements reserved by the subsequent budget-$2$ LOSS leaf:
\begin{equation}
 E_{P_0}^\D(W)=1+2=3.
 \label{eq:running-full-exposure}
\end{equation}
Consequently the conservative ordinary touched calculation reads
\[
 \cnt_\D(W,P_0)+E_{P_0}^\D(W)=3+3=6.
\]
If this full clock is used at an ordinary zone node, all three empties of $W$
belong to its touched term.  The gate clock introduced in \cref{def:D20a}
instead has the verified value
\[
 Q_{P_0}^\D(W)=\max\{1,0+2\}=2,
\]
because the exact copied reply $h$ lies outside $W$.  The resulting fixed-
window total is $3+2=5$.

The second query is
\[
 W'=\{(q,0):5\le q\le10\}.
\]
It contains one root defender stone, at $(8,0)$, and the copied reply
$h=(5,0)$ lies in it.  The branch-coherent gate recurrence therefore gives
\[
 Q_{P_0}^\D(W')=\max\{1,1[h\in W']+2\}=3.
\]
The three units are attained by $h$ followed by the LOSS-remainder placements
$(6,0)$ and $(7,0)$.  The gate proof in \cref{sec:gates} explains why the
first value may be debited and the second may not.

\subsection{A short static cover}

The ranked zone may sometimes be covered by a position-local set.  This is a
restricted theorem, not a general replacement for certificate obligations.

\begin{theorem}[Static-set coverage for short local budgets]
\label{thm:T5}
\upshape
Suppose $B(N)\le3$, at most three remaining attacker placements are
threat-creating, every named witness-empty role lies in an attacker-touched
window, and $\Zseed(N)=\varnothing$.  Then
\[
 r_3(P_N)\cup
 \{\text{empties of attacker-touched alive windows at }P_N\},
\]
where $r_3(P_N)$ is the set of legal empties within distance $3$ of a stone,
covers $Z(N)$.  Without all of these hypotheses, the displayed set is a
heuristic candidate generator rather than a certified zone.
\end{theorem}

\begin{plainterms}
Under these simultaneous hypotheses, two easy local sets already contain
every mandatory zone term.  Removing the short-budget, seed, or witness
hypothesis can restore remote obligations, so this is not a general
replacement for the ranked zone.
\end{plainterms}

\begin{proof}[Proof sketch]
Every exposure is at most $B(N)\le3$, so the virgin term is empty.  A touched
window in the completion term has at least $6-E_N^\D(W)\ge3$ defender stones,
and \cref{lem:L1} places each of its empties in $r_3(P_N)$.  The first-three
containment of \cref{lem:L10} covers the stated designated moves, while the
witness hypothesis covers leaf roles; the seed term is empty by assumption.
See \cite{hexobot-companions}.
\end{proof}

\subsection{The extendable-hit kernel}

At a fresh equal-position node, the current attacker threats can replace the
certificate's future obligations for a short forced region.  The remaining
reply must not merely hit one threat; it must leave a hitting set that fits in
the rest of the defender budget.

For the current attacker-threat empty-set family $\mathcal F_N$, write
\[
 \mathcal F_N\setminus d=\{E\in\mathcal F_N:d\notin E\},
 \qquad
 K_b(N)=\{d\in\Legal(P_N):
               \tau(\mathcal F_N\setminus d)\le b-1\}.
\]

\begin{theorem}[Extendable-hit kernel]
\label{thm:T6}
\upshape
A kernel-governed region is entered at a certificate node where the real and
ghost positions are equal.  It uses the following rule until a typed terminal
or an $\mhs>b$ handoff.  At every internal AND node in the region, require the
explicit defender check $\neg\ownwinnow$ and
\[
 \mhs(P_N)=\tau(\mathcal F_N)\le b,
\]
and search exactly $K_b(N)$.  No original obligation or core term is required.
If $\mhs<b$, then $K_b(N)=\Legal(P_N)$ and there is no pruning.  If
$\mhs=b$, the kernel weakly refines the current hitting-cell set and can
strictly reduce it; equality can occur.  A node with $\mhs>b$ ends the kernel
region.  It must be a valid LOSS leaf or hand off to an existing nonempty
sound subtree, below which the core-free claim is not asserted without a new
equal-position entry.
\end{theorem}

\begin{plainterms}
The kernel keeps only a first hit that can be extended to a complete defense
of the current threat family within the remaining placements.  When the
family needs fewer hits than the budget, every legal reply is extendable and
the rule saves nothing.  When the pressure exceeds the budget, the node must
use a LOSS contract or return to a separately certified subtree.
\end{plainterms}

\begin{proof}[Proof sketch]
If $\tau<b$, deleting the sets hit by any legal $d$ leaves transversal number
at most $b-1$, while if $\tau=b$ the first member of a minimum transversal is
legal by \cref{thm:T1} and lies in the nonempty kernel.  At the first reply
outside the kernel the positions still agree, so the original subtree may be
abandoned: for $b=2$ the residual family gives a budget-$1$ LOSS, and for
$b=1$ one threat survives for the following budget-$2$ attacker turn.  The
explicit immediate-win exclusion bounds every defender-alive window by three
stones at $b=2$ or four at $b=1$, so the defender cannot complete first.
Following a minimum transversal through the searched kernels kills all
current threats and leaves the original horizon enough time for the auxiliary
two-placement refutation.  If $\tau>b$, the kernel is empty, which explains
the required handoff.  See \cite{hexobot-companions}.
\end{proof}

\subsection{Ranked relevance closure}

The mandatory closure is the ranked zone itself.  A search implementation may
add three finite local heuristics without making them verifier premises.

\begin{definition}[Ranked relevance closure]
\label{def:D13}
Define
\[
 \Rcert(\mathcal C,N)=
 \Zdir(N)\cup\Zseed(N)\cup\Ztouch(N)\cup\Zvirgin(N).
\]
The solver may search the finite superset
\[
 \Rsearch(\mathcal C,N)=\Rcert(\mathcal C,N)
 \cup\operatorname{hitting}(P_N)\cup\mathcal A(P_N)\cup r_3(P_N),
\]
where $\operatorname{hitting}(P)$ is the union of the empty sets of current
attacker threats, $\mathcal A(P)$ is the set of empties in attacker-touched
alive windows, and $r_3(P)$ is the legal radius-$3$ neighborhood of the
current stones.  The last three terms are heuristics, not hypotheses of
\cref{thm:T3}.  The mandatory terms are finite: obligations come from a
finite certificate, touched windows from a finite position, and each virgin
query examines only windows within its bounded exposure radius of a finite
legal candidate.
\end{definition}

\begin{plainterms}
$\Rcert$ is the verifier boundary; every cell in it must be covered.
$\Rsearch$ may be larger to help discovery, but its three added sets are not
soundness premises.  Omitting a heuristic cell can lose a proof, whereas
omitting a mandatory cell can make an accepted proof unsound.
\end{plainterms}

The next theorem turns this computed closure into a verifier condition.

\begin{theorem}[Ranked-closure coverage]
\label{thm:T7}
\upshape
At every ordinary zone-governed AND node, any independently nonempty
$S(N)\supseteq\Rcert(\mathcal C,N)$ satisfies \cref{zone:Z2} and
\hyperref[zone:Z5prime]{clause \textup{(Z5$'$)}}.  Dismissal soundness follows
from \cref{thm:T3} when the D9 grammar, the D14--D16 clock checks, and
\cref{zone:Z4} also hold.  In constructing $\Rcert$, the verifier unions roles
over all reachable descendants, includes every designated attacker move
including OR-COMPLETION, treats WIN and LOSS continuation cells as leaf-entry
witness-empty roles, takes the maximum rank over live roles sharing a cell,
and forms no band after the $r=0$ deadline check.  Defender-completion windows
remain on the $E^\D$ clock or the conservative $B$ clock.
\end{theorem}

\begin{plainterms}
Covering $\Rcert$ is sufficient only together with an independent searched
fallback, the certificate grammar, the clock checks, and attacker legality.
The statement validates a supplied certificate; it does not promise that a
heuristic candidate generator will find one.
\end{plainterms}

\begin{proof}[Proof sketch]
The four terms in $\Rcert$ are exactly the four sets in \cref{def:D11}, so
coverage supplies Z2 and Z5$'$.  The stated role-union and deadline rules are
the hypotheses needed to compute those sets rather than approximations to
them.  With Z4 and the certificate clocks checked, \cref{thm:T4,thm:T3}
apply.  See \cite{hexobot-companions}.
\end{proof}

\subsection{Branch-indexed substitution}

The global zone protects the union of every reachable branch.  A dismissal
may instead select one searched child and verify a smaller envelope over all
descendants of that child.  The current real and ghost placements are already
different, so every selected clock must include that transition.

\begin{definition}[Transition-inclusive substitution envelope]
\label{def:D17}
At an AND node $N$, a ghost-legal dismissal $d$ may name a searched substitute
\[
 \varphi_N(d)=s\in S(N)
\]
with exact child $C_s$.  The node also declares a nonempty fallback
$F(N)\subseteq S(N)$ independently of these annotations.  In particular,
$S(N)$ is not defined circularly as the replies for which no safe substitute
exists.  A valid selected envelope has the following data and checks.
\begin{enumerate}[label=\textup{(\arabic*)},leftmargin=10mm]
  \item Its transition budget is
  $\widehat B(N,d,s)=1+B(C_s)$.  It includes the current real/ghost defender
  ply and every reachable LOSS remainder.

  \item Its obligation set is the union of the roles at all reachable
  descendants of $C_s$, not the roles of one leaf or one selected
  continuation.

  \item The real cell $d$ avoids every selected obligation cell and every
  transition-dangerous completion empty.

  \item If a selected child obligation $y$ has rank $r_{C_s}(y)$ and
  $y\notin\Legal(P_N)\cup\Stones(P_N)$, direct occupation $d=y$ is forbidden
  and the parent seed test is
  \[
    \dist(d,y)>8r_{C_s}(y)
      =8\bigl((1+r_{C_s}(y))-1\bigr).
  \]

  \item For every defender-alive touched window $W$ with
  $\cnt_\D(W,P_N)\ge1$, the $\widehat B$ test forbids
  $d\in E(W,P_N)$ whenever
  \[
    \cnt_\D(W,P_N)+1+B(C_s)\ge6.
  \]
  For an all-empty window it forbids a seed satisfying
  $\dist(d,W)\le8(\widehat B-6)$ when $\widehat B\ge6$.  The exact-exposure
  form instead sets
  \[
    \widehat E^\D(W)=1+E_{C_s}^\D(W)
  \]
  and uses $\cnt_\D(W,P_N)+\widehat E^\D(W)\ge6$ for a touched fill and
  $\dist(d,W)\le8(\widehat E^\D(W)-6)$, with
  $\widehat E^\D(W)\ge6$, for an all-empty window.  Both touched and virgin
  clauses are mandatory for the selected clock.

  \item The fallback $F(N)$, and hence $S(N)$, remains nonempty independently
  of whether any dismissal has a valid substitute.

  \item For a ghost-legal A3 dismissal, the ghost plays $\varphi_N(d)$.  Case
  A2 may use any searched filler because it creates no new $X$-stone.  A
  ghost-illegal A3 dismissal inherits the envelope of the earlier ghost-legal
  seed that supports its real legality and may use any searched filler in the
  current reachable subtree.  It does not start a child-only clock.

  \item The envelope protects every reachable LOSS witness-empty role through
  leaf entry, while $\widehat B$ or $\widehat E^\D$ counts the leaf's $b$
  placements for defender completion and defender-own-win exclusion.
\end{enumerate}
The default-child form declares $f(N)\in F(N)$ and uses $C_{f(N)}$ for every
ghost-legal dismissal.  It must satisfy the same transition-inclusive checks.
Replacing a whole-tree role union by a child union without these checks is
insufficient.
\end{definition}

\begin{plainterms}
The real defender may play $d$ while the ghost follows the searched reply
$s$, so the verifier protects the chosen child's entire future against that
new mismatch.  The leading unit charges the current transition in both the
role and window clocks.  A child-only calculation would miss a legality link
or a defender fill created on this ply.
\end{plainterms}

The envelope is useful only if it preserves the main dismissal conclusion.

\begin{theorem}[Branch-indexed substitution soundness]
\label{thm:T9}
\upshape
The conclusion of \cref{thm:T3} remains valid if an AND node's global
\cref{def:D11} dismissal tests are replaced, dismissal by dismissal, by a
valid \cref{def:D17} envelope.  The same holds for the default-child form.
\end{theorem}

\begin{plainterms}
A valid envelope permits branch-specific substitution, not an arbitrary reply
that happens to hit the same threat.  The descendant union and
transition-inclusive clocks preserve all three causal channels.  If any check
fails, the node is not certified by this theorem.
\end{plainterms}

\begin{proof}[Proof sketch]
The real reply and ghost substitute consume the same defender ply, while the
transition cell and the leading $1$ account for occupancy, legality-frontier,
and mask differences at that step.  Every later ghost node lies among the
declared descendants of $C_s$; nested selected envelopes preserve avoidance
for earlier $X$-stones, and a ghost-illegal chain inherits the envelope of its
first ghost-legal seed.  The touched and virgin tests give the fixed-window
argument through every reachable LOSS remainder.  Z4, leaf-mask transfer, and
the terminal cases of \cref{thm:T3} are unchanged.  See
\cite{hexobot-companions}.
\end{proof}

The leading $1$ is needed in both hazard channels.  For occupancy, a legal
real reply at distance $8$ from a future ghost-illegal obligation can legalize
that obligation for the remaining defender placement; a child rank of one
would give radius zero without the transition and radius eight with it.  For
completion, two defender stones in a window, a real transition into that
window, and three later fills total six even though the child-only count is
$2+3=5$.

The ordinary completion zone also derives the internal immediate-win
diagnostic.

\begin{lemma}[Internal own-win diagnostic]
\label{lem:L14}
Under the completion-zone requirement of \cref{def:D11} and the ban in
\cref{def:D9} on defender-terminal edges, an internal AND node cannot contain
a defender-alive count-$5$ window, or a count-$4$ window when the defender
budget is $2$.
\end{lemma}

\begin{proof}[Proof sketch]
A count-$5$ touched window has exposure at least one, so its legal last empty
is searched and its exact child would be defender-terminal.  At budget $2$, a
count-$4$ touched window has exposure at least two, so both empties are
searched; either first fill reaches a budget-$1$ count-$5$ child.  Such a child
cannot be a LOSS leaf because its mandatory immediate-win check fails, and an
internal child would again search the terminal last empty.  See
\cite{hexobot-companions}.
\end{proof}

The check remains mandatory at LOSS leaves and in the kernel theorem, whose
governed region has no ordinary completion guard.

\subsection{Finite certificate DAGs}

Identical certificate subtrees may be shared, provided every route into a
shared node agrees on its exact state and clock.

\begin{definition}[Finite acyclic certificate DAG]
\label{def:D18}
A D9 certificate may be a finite rooted acyclic graph if every shared node
has one exact D9 label: the same position, mover and budget, designated OR
action or AND searched-successor map, leaf witness data, and one consistent
path clock.  Obligations at a DAG node are unions over all reachable
descendants.  Local budgets, ranks, and exposures satisfy their inequalities
on every outgoing edge.  Coupling histories $X$, $Y$, and $\widehat X$ remain
path-local.
\end{definition}

\begin{plainterms}
A DAG may share a certificate state only when every incoming route agrees on
the state, action, witnesses, and clocks.  The stored node is shared, but the
real--ghost difference history remains attached to a path.  Merging
inconsistent labels would no longer describe one certificate node.
\end{plainterms}

Soundness follows by expanding the shared representation back into a tree.

\begin{theorem}[DAG unfolding]
\label{thm:T10}
\upshape
Every \cref{def:D18} DAG satisfying the ordinary \cref{def:D11} rules or the
substitution rules of \cref{def:D17} has the corresponding soundness
conclusion of \cref{thm:T3}.
\end{theorem}

\begin{plainterms}
The DAG saves storage but proves no new play.  Unfolding makes one copy of a
shared node for each incoming path and leaves every verified label unchanged.
A cycle or inconsistent clock would prevent this finite tree reduction.
\end{plainterms}

\begin{proof}[Proof sketch]
A finite acyclic graph has finitely many root-to-node paths.  Unfold one copy
of each shared node along each path; the exact labels and consistent clocks
give a finite D9 tree with unchanged transitions, terminal data, clock
inequalities, and descendant-role unions.  Apply \cref{thm:T3} or
\cref{thm:T9}; the unfolded coupling histories are precisely the required
path-local histories.  See \cite{hexobot-companions}.
\end{proof}

At verification time, an ordinary node must establish $\Rcert$ coverage and
Z4, or validate every substitution annotation, while retaining an independent
fallback.  A failed check returns \textsc{unknown}.  Heuristic search trimming
can affect completeness, but a certificate that passes these checks remains
sound.

\subsection{Local domination patterns}
\label{sec:domination}

The zone theorems compare a real reply with a ghost certificate path.  Four
additional local rules compare two defender replies directly.  Their proofs
use causal strategy transfer and check the same occupancy, window-mask, and
legality-frontier channels.

For this subsection, let
\[
 \Lambda(P)=\bigcup_{s\in\Stones(P)}B_8(s)
\]
be the supported cells, and let $\Omega(x)$ be the 18 windows containing
$x$.  For legal defender replies $a,b$ at the same position, write
$b\preceq_n a$ when, for every defender continuation strategy after $b$,
there is one after $a$ such that, for every attacker challenge after $a$,
there is an attacker continuation after $b$ satisfying
\[
 u_\D\bigl(\operatorname{Out}_n(P+a,\sigma_a,\tau_a)\bigr)
 \ge
 u_\D\bigl(\operatorname{Out}_n(P+b,\sigma_b,\tau_b)\bigr),
\]
where $u_\D$ assigns $1$ to a defender win, $0$ to no terminal event through
the horizon, and $-1$ to an attacker win.  Thus an attacker win against the
searched reply $a$ transfers to the omitted reply $b$.

A dead empty cell supplies no new radius-$8$ legality frontier.

\begin{lemma}[A dead empty has no legality frontier]
\label{lem:Dom-L7}
Let $b$ be empty and dead in a nonterminal post-opening position $P$, meaning
that every window in $\Omega(b)$ is dead.  Then
\[
 B_8(b)\subseteq\Lambda(P).
\]
In particular, $b$ is legal, and placing at $b$ adds no unsupported
cell to the legal store.
\end{lemma}

\begin{proof}[Proof sketch]
For each directed axial unit vector $u$, the endpoint window
$\{b,b+u,\ldots,b+5u\}$ is dead and contains an old stone
$s_u=b+k_u u$ with $1\le k_u\le5$.  A cell $z\in B_8(b)$ lies in a
$60^\circ$ sector between adjacent directions; choosing the stone on either
boundary ray gives $\dist(z,s_u)\le8$.  Hence every point of $B_8(b)$ was
already supported.  See \cite{hexobot-companions}.
\end{proof}

Deadness of the omitted cell does not make an arbitrary searched substitute
safe.  The substitute must also be frontier-inert unless it wins immediately.

\begin{proposition}[Frontier-certified dead-cell dismissal]
\label{prop:Dom-P1}
Let $P$ be a nonterminal post-opening position with $\D$ to place, and let
$a\ne b$ be legal empty cells.  Assume every window in $\Omega(b)$ is already
dead.  Unless $P+a$ is an immediate defender win, also assume
\[
 \Lambda(P)\cup B_8(a)=\Lambda(P)\cup B_8(b).
\]
Then $b\preceq_n a$ for every finite $n$.  By \cref{lem:Dom-L7}, the displayed
support equality is equivalent here to $B_8(a)\subseteq\Lambda(P)$.
\end{proposition}

\begin{proof}[Proof sketch]
Compare $P+a$ and $P+b$ by transposing the two special coordinates until the
counterpart is used.  The support equality makes the transposition legal at
every paired prefix.  All windows through $b$ are permanently dead; the other
changed windows either agree or change ownership in the defender-favouring
direction.  Per-placement terminal checks stop the transfer at the first win,
and an immediate defender win after $a$ is already maximal.  This gives an
all-depth alternating simulation and hence the stated domination.  See
\cite{hexobot-companions}.
\end{proof}

Two hits of one count-$4$ threat are interchangeable only when their other
incident windows and their successor supports are made inert.

\begin{proposition}[Dead-spoke interchangeable hits]
\label{prop:Dom-P2}
Let $P$ be a nonterminal post-opening position with $\D$ to place.  Let $W$
be an attacker-alive window with exactly four attacker stones and exactly two
empty cells $x\ne y$.  Both cells are legal.  Assume every window in
\[
 (\Omega(x)\cup\Omega(y))\setminus\{W\}
\]
is already dead in $P$, and assume
\[
 \Lambda(P)\cup B_8(x)=\Lambda(P)\cup B_8(y).
\]
Then the two replies are mutually outcome-dominating at every finite horizon:
\[
 x\preceq_n y,
 \qquad y\preceq_n x.
\]
A solver may retain one canonical member of $\{x,y\}$ at that defender node.
\end{proposition}

\begin{proof}[Proof sketch]
Either defender hit kills $W$, and the dead-spoke premise then makes every
window through either special cell permanently dead.  Transposing $x$ and
$y$ preserves occupancy choices, the support equality preserves legal moves
in both directions, and all masks outside the dead union agree.  This is an
outcome-preserving bisimulation at every depth.  Equality of window count
profiles alone would not preserve labelled empties or their intersection
hypergraph.  See \cite{hexobot-companions}.
\end{proof}

The two placements of one turn can be stored as an unordered pair only if
both orders were legal before the first placement and neither singleton prefix
wins.

\begin{proposition}[Nonwinning-prefix turn commutation]
\label{prop:Dom-P3}
Let $P$ be a nonterminal \textup{FirstStone} position for either player $X$.
Let $p\ne q$ both be legal in $P$, and suppose neither transition $P+p$ nor
$P+q$ is terminal.  Then both two-placement traces $p;q$ and $q;p$ are legal
and have the same qualitative rule outcome after the second placement.  If
nonterminal, they have identical occupancy, labelled window masks, legal
support, player to move, and \textup{FirstStone} phase.  If terminal, both
terminate on the second placement with winner $X$ and the same absolute
placement count.
\end{proposition}

\begin{proof}[Proof sketch]
Initial legality makes either second cell remain legal, and the final support
$\Lambda(P)\cup B_8(p)\cup B_8(q)$ is independent of order.  Adding the two
owner bits to every window mask commutes, including on a window containing
both cells.  The two singleton-nonterminal premises exclude a first-placement
win, so any joint completion occurs on the second placement in both orders.
The live engine states may retain different order-sensitive history metadata;
the proposition equates rule outcomes, not serialized state fields.  See
\cite{hexobot-companions}.
\end{proof}
