\section{Mechanized validation}
\label{sec:validation}

The main soundness results are mathematical proofs.  Some auxiliary claims
also have finite machine checks.  Other computations measure a search
generator or test a local rule on sampled positions.  These roles are
different.  This section records the input class, the published count, and
the conclusion supported by each run.  The scripts and full traces are
described in the repository companion documents \cite{hexobot-companions}.

\subsection{Search-harness runs}

The compact search model reproduces the engine's coordinates, legality,
phase machine, per-placement terminal check, and threat masks.  Random replay
found matching legal sets and terminal decisions after each placement.  No
aggregate concordance count is published, so this remains implementation
evidence rather than an exhaustive equivalence proof.

On 400 random positions, the largest observed distances were two for an
immediate completion or threat hit and three for threat creation.  A separate
universe measurement found means of about 302 legal cells, 121 radius-three
cells, and 80 radius-two cells.  The radius-three plus attacker-touched predecessor
generator later measured about 147 cells against the roughly 302-cell legal
baseline.  These are measurements, not verifier constants.

The engineered junction position supplies a negative control.  Restricting
the defender to the radius-two set produced an attacker win after 2,374
uncapped search nodes.  The run proved all 217 turns of the form
``mandatory pin hit plus radius-two spare'' losing under that restriction.
The omitted distance-three junction reply, paired with the same mandatory
hit, was not refuted by an uncapped search through depth eight.  The
comparison establishes that radius two is an unsound defender restriction.
It does not establish that the junction reply is a full-game defender win.
The defender-side survival probe stops at the first unrefuted child, which is
valid for non-loss at the tested horizon but not for a LOSS proof.

A comparison on 109 attackable positions found no genuine radius-two versus
full-set divergence.  One node-cap artifact was re-adjudicated at two million
nodes and gave the same \textsc{win} verdict in both variants.  A sparse
failure mode still invalidates a universal pruning rule.

On the engineered junction, the predecessor closure contained the junction
cell and the defender survived after 1,396 uncapped nodes.  A bounded run on
91 attackable positions produced 52 restricted \textsc{win} claims, zero
full-set divergences at the matched horizon, and a time of 139 seconds.  The
radius-two controls show that the harness can expose an unsound restriction.
The closure run remains search evidence; it does not prove \cref{thm:T7} or
replace \cref{thm:T3,thm:T4}.

\begin{table}[tbp]
  \centering
  \caption{Published search-harness counts and their scope.}
  \label{tab:search-validation}
  \small
  \renewcommand{\arraystretch}{1.08}
  \begin{tabular}{@{}p{.22\textwidth}p{.25\textwidth}p{.43\textwidth}@{}}
    \toprule
    Run & Published count & Supported conclusion \\
    \midrule
    Local-distance sample & 400 positions; maxima 2 and 3 &
      Confirms the predicted local distances on the sample only. \\
    Radius-two junction & 2,374 nodes; 217 restricted turns &
      Gives a concrete false-pruning control for radius two. \\
    Random radius-two comparison & 109 positions; 0 genuine divergences &
      Measures rarity in the sample; does not restore soundness. \\
    Closure junction & 1,396 uncapped nodes &
      Confirms that the tested closure retains the known junction defense. \\
    Closure random comparison & 91 positions; 52 claims; 0 divergences;
      139 seconds &
      Tests the generator at matched horizons; does not prove ranked-zone
      coverage. \\
    \bottomrule
  \end{tabular}
\end{table}

\subsection{Domination spot checks}

The domination rules in \cref{sec:computable-zones} have analytic proofs.
Two finite randomized checks exercise their most error-prone rule channels.
The turn-commutation check generated 2,997 legal same-mover pairs after
filtering out every singleton-winning prefix.  It found zero differences in
the final occupancy, masks, legal support, next phase, or terminal winner.
The dead-cell check generated 400 adversarial configurations.  The stones
that make the focal cell dead were placed at large permitted offsets.  It
found zero violations of the radius-eight frontier containment in
\cref{lem:Dom-L7}.  The worst cell in the focal radius-eight ball was at
distance six from an existing stone, leaving a distance margin of two.

These checks are not exhaustive proofs of the domination patterns.  The
companion gives finite quotient specifications for MV-P1, MV-P2, and MV-P3,
including 289 marked local geometries for MV-P1 and 81 for MV-P3 after the
stated local-incidence quotient.  It also counts
$942{,}779{,}391{,}290$ dead-star colourings up to the dihedral symmetry in
the MV-P1 regression space.  Those numbers describe prospective enumeration
spaces.  The full quotient enumerations have not been run.  In particular,
the 2,997-pair and 400-position checks do not justify dropping the
frontier-inertness premise of P1, the dead-spoke and support premises of P2,
or either singleton premise of P3.

\subsection{Four finite-check campaigns}

Four focused campaigns accompany the forced-hit, tightness, pairing, and
potential results.  The review log records that the repaired
assertion-enabled checkers exit successfully.  Only the pairing and potential
reports publish exhaustive state counts.  The tightness and forced-hit
reports publish the checked predicates but no aggregate assertion count.

\paragraph{Threshold pairing.}
The pairing checker works on the twelve-cell quotient of the period lattice
in \cref{thm:T8.2}.  Its exact-cover instance has 36 rows and 18 columns: six
line-cycle constraints and twelve endpoint constraints.  Direct phase space
has $6^6=46{,}656$ assignments.  An MRV Algorithm~X traversal exhausted the
instance in 419 recursive states and found 120 labelled solutions.  A
separate direct iteration over all 46,656 phase tuples also found 120.  The
displayed six-pair construction occurs in both enumerations.

The checker verifies six pair orbits and twelve quotient cell orbits.  The
cell-incidence histogram is $\{1:12\}$.  It verifies 36 quotient
start-and-axis window orbits, with contained-pair histogram $\{1:36\}$.
As a redundant finite-patch test, it checks all 3,721 cells in
$[-30,30]^2$ and all 11,163 length-seven windows starting there.  The two
patch histograms are $\{1:3721\}$ and $\{1:11163\}$.  Translation by the
period lattice makes the quotient checks exhaustive for the infinite
construction.  This computation establishes the stated matching and
coverage properties and exhausts the six phase variables for that quotient.
The index-minimality argument is analytic.  The computation does not produce
a pairing for length six, does not replace the density proof of
\cref{thm:T8}, and does not rule out position-dependent partial pairings.

\paragraph{Exact-surd greedy refutation.}
The potential checker represents each value as an integer pair for
$(a+b\sqrt3)/27$ and compares pairs without floating point.  It includes
every exact greedy maximizer along the fixed attacker continuation in
\cref{thm:ES-greedy-refutation}.  The successive input/output state counts
for the eight defender placements are
\[
  1/4,\quad 4/4,\quad 4/12,\quad 12/124,\quad
  124/242,\quad 242/124,\quad 124/124,\quad 124/124.
\]
State deduplication merges only identical coloured boards at the same phase.
At the last defender placement, 106 states have unique second maximum
$(-3,1)$ and 18 have unique second maximum $(-2,-1)$.  All 124 branches omit
the attacker's winning cell.  The checker verifies legality, every
nonterminal prefix, the exact alive-window profiles, every maximum and tie,
and one terminal window on all 124 branches.  It also constructs a legal
39-placement engine history for the enlarged reachable witness and replays
the same full greedy tree there.

There is no randomness, beam, or depth cutoff in this enumeration.  It proves
that the named fixed attacker continuation defeats every dynamic-greedy
tie-breaking on the stated position.  It does not prove that every defender
strategy loses from every position with $\Phi<1$.  That raw existential
question remains open.  It also does not weaken the fixed-family and
finite-region sufficiency theorems, whose target families do not renew in the
same way.

\paragraph{Tightness checks.}
The tightness checker publishes the following summary.  The triangle has
three threats, transversal number two, and minimum witness size three.  The
five-cycle has five threats, transversal number three, and minimum witness
size five.  The T6 diagnostic has $K_1=\{k\}$ while the omitted cell is an
immediate defender completion.  The OR-COMPLETION and leaf-entry deadline
roles pass their protection checks.  The exact-rank, virgin-radius,
touched-equality, and T5 local distance/count traces pass.  The report then
states that all tightness checks passed.  It publishes no aggregate number of
assertions or configurations.

The checker establishes the listed finite mask, transversal, deadline, and
arithmetic facts.  The remaining coordinate families and framework-scope
arguments in the tightness report are prose proofs.  In particular, a
checked equality for one fixed window does not create a complete
countercertificate for the full union of zone terms.  The OPEN and relative
classifications in \cref{sec:tightness} therefore remain part of each result.

\paragraph{Forced-hit review.}
The forced-hit report gives an eight-stage finite verifier procedure.  It
recomputes each named threat mask and empty set; checks the transversal number
and the defender non-win diagnostic; enumerates the kernel; checks exact
successors and the ban on kernel substitution; propagates checkpoint roles;
computes the least $f$ and queried $Q^{\D}$ values in reverse topological
order; enumerates the four debited zone terms at ordinary nodes; and checks
every escape deadline.  The 15-plus-15 sharpness position separately checks
all 18 incident windows per stone, nonterminality, and the stated maximum
defender count.  The report and review log record a successful
assertion-enabled checker run, but they publish no aggregate count of
assertions, queried windows, or certificate nodes.

The machine run is a consistency check on finite data and recurrences.  The
proof of \cref{thm:T11} is analytic and has status \textsc{proven}, not
\textsc{proven-mech}.  The run does not prove a target-independent reduction
of $B$, a zero-cost substituted hit, automatic net shrinkage of the final
searched set, or a debit for slack pressure.  Those nonclaims are part of the
PARTIAL boundary of the gate calculus.

\subsection{Validation boundary}

The validation layers support different statements.  Exhaustive quotient
enumeration proves the finite pairing construction.  Exact branch
enumeration proves the all-ties greedy counterexample.  The tightness and
forced-hit checkers confirm named finite calculations while their main
claims remain analytic or scope-qualified.  Randomized domination checks
test implementations of proven local rules.  Search-harness runs measure a
candidate generator and supply negative controls.  None of these runs allows
a verifier to accept an omitted defender reply without the hypotheses of the
theorem used at that node.
